Follow-up Samples

### 2016-12-05 submission/manuscript_outline_20161024.Rmd
update abstract again

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this study, we assessed whether a model built with bacterial microbiome data could accurately predict adenoma or carcinoma and adjust positive probability for lesion (adenoma or carcinoma) after surgical treatment.  This model was tested on a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  : For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```).  The top 5 most important operational taxonomic units (OTUs) for prediction of lesion classified to ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  In the test set, the prediction of lesion for the initial samples had a sensitivity of ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) while follow up it was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, this model could accurately predict those with an adenoma or carcinoma and lower probability of having a lesion based on follow up samples after treatment.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease.  Previous research suggests that the bacterial microbiome can be used as a biomarker for colorectal cancer (CRC).  These reports have mostly focused on investigating how the bacterial microbiome is used at a single point in time to predict disease.  In this study, we assessed whether a model built with bacterial microbiome data could accurately predict adenoma or carcinoma and adjust positive probability for lesion (adenoma or carcinoma) after surgical treatment.  This model was tested on a ```r sum(initial_cm$ref_yes)``` person group that included samples before and after treatment to allow for the assessment of how the model adjusts risk after treatment. The model chosen had a 10-fold cross validated AUC of ```r format(model_test_AUC_summary["middle", "ROC_cv"], digits = 3)``` and a test set AUC of ```r format(model_test_AUC_summary["middle", "AUC"], digits = 3)```.  : For the follow up samples our Random Forest model significantly decreased the positive probability of lesion compared to the initial samples for both adenoma (P-value = ```r format(IF_model_wilcox_paired_summary['all_adenoma', 'BH_correction'], scientific = TRUE, digits = 3)```) and carcinoma (P-value = ```r format(IF_model_wilcox_paired_summary['carcinoma_only', 'BH_correction'], scientific = TRUE, digits = 3)```).  The top 5 most important operational taxonomic units (OTUs) for prediction of lesion classified to ```r paste(rf_lowest_ID[1], " (", names(rf_lowest_ID)[1], ")", sep = "")```,  ```r paste(rf_lowest_ID[2], " (", names(rf_lowest_ID)[2], ")", sep = "")```, ```r paste(rf_lowest_ID[3], " (", names(rf_lowest_ID)[3], ")", sep = "")```, ```r paste(rf_lowest_ID[4], " (", names(rf_lowest_ID)[4], ")", sep = "")```, and ```r paste(rf_lowest_ID[5], " (", names(rf_lowest_ID)[5], ")", sep = "")```.  In the test set, the prediction of lesion for the initial samples had a sensitivity of ```r format(confusion_summary["Sensitivity", "initial"]*100, digits = 3)```% (```r paste(initial_cm["pred_yes", "ref_yes"], "/", sum(initial_cm$ref_yes), sep = "")```) while follow up it was ```r format(confusion_summary["Sensitivity", "followup"]*100, digits = 3)```% (```r paste(followup_cm["pred_yes", "ref_yes"], "/", sum(followup_cm$ref_yes))```).  Overall, our model accurately predicted those with an adenoma or carcinoma and following treatment it also decreased the positive probability of having a lesion.

### 2016-12-05 submission/manuscript_outline_20161024.tex
update abstract again

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a \texttt{66} person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model chosen had a 10-fold cross validated AUC
of \texttt{0.819} and a test set AUC of \texttt{0.815}. : For the follow
up samples our Random Forest model significantly decreased the positive
probability of lesion compared to the initial samples for both adenoma
(P-value = \texttt{8.48e-07}) and carcinoma (P-value =
\texttt{5.96e-08}). The top 5 most important operational taxonomic units
(OTUs) for prediction of lesion classified to
\texttt{Lachnospiraceae\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Ruminococcaceae\ (Otu000020)},
\texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. In the test set, the prediction of
lesion for the initial samples had a sensitivity of \texttt{98.5}\%
(\texttt{65/66}) while follow up it was \texttt{100}\%
(\texttt{1\ /\ 1}). Overall, this model could accurately predict those
with an adenoma or carcinoma and lower probability of having a lesion
based on follow up samples after treatment.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a \texttt{66} person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model chosen had a 10-fold cross validated AUC
of \texttt{0.819} and a test set AUC of \texttt{0.815}. : For the follow
up samples our Random Forest model significantly decreased the positive
probability of lesion compared to the initial samples for both adenoma
(P-value = \texttt{8.48e-07}) and carcinoma (P-value =
\texttt{5.96e-08}). The top 5 most important operational taxonomic units
(OTUs) for prediction of lesion classified to
\texttt{Lachnospiraceae\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Ruminococcaceae\ (Otu000020)},
\texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. In the test set, the prediction of
lesion for the initial samples had a sensitivity of \texttt{98.5}\%
(\texttt{65/66}) while follow up it was \texttt{100}\%
(\texttt{1\ /\ 1}). Overall, our model accurately predicted those with
an adenoma or carcinoma and following treatment it also decreased the
positive probability of having a lesion.
